Inside Out: Externalizing Assumptions in Data Analysis as Validation Checks

### 2025-07-21 index.tex
revision done

\phantomsection\label{cell-fig-step-count}
\begin{figure}[H]

@@

\phantomsection\label{cell-fig-metric-calc}
\begin{figure}[H]

@@

While checks may score high on predictive accuracy, they may be
correlated and less useful to provide actions to act on the analysis.
This could happen if a set of checks are all tangentially related to the
cause of the unexpected results, but none addresses the root cause. The
independence score in Figure~\ref{fig-metric-calc} is defined as 1 minus
total correlation per observation.

Total correlation is a generalization of mutual information that
measures the amount of shared information across multiple variables,
based on Kullback-Leibler distance:

\[C(X_1, X_2, \cdots, X_k) = \sum_{x_1} \sum_{x_2} \cdots \sum_{x_k} p(x_1, x_2, \cdots, x_k) \log \frac{p(x_1, x_2, \cdots, x_k)}{p(x_1)p(x_2) \cdots p(x_k)},\]

where \(p(x_1, x_2, \cdots, x_k)\) is the joint distribution of the
predictors and \(p(\cdot)\) is the individual distribution of the
predictors. A high total correlation value indicates overlapping of
information among the analysis validation checks, while a low value
suggests that the checks are independent and provides unique information
to diagnose the unexpected outcome. To standardize this measure, the
total correlation \emph{per observation} is used in calculating the
independence score.

@@

\phantomsection\label{cell-fig-logic-reg}
\begin{figure}[H]

@@

\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}{>{\raggedright\arraybackslash}p{17em}rrrrr}
\toprule
Checks & Precision & Recall & Independence & Harmonic & Arithmetic\\
\midrule
Check 1: q(step, 0.6) > 10000 & 0.319 & 0.575 & 1.000 & 0.511 & 0.631\\
Check 2: q(step, 0.4) < 8000 & 0.264 & 0.613 & 1.000 & 0.467 & 0.626\\
Check 3: sd(step) > 2500 & 0.153 & 0.289 & 1.000 & 0.273 & 0.481\\
Logic regression: (check 1 OR check 2) AND (not check 3) & 0.861 & 0.431 & 0.998 & 0.669 & 0.763\\
Regression tree & 0.542 & 0.780 & 1.000 & 0.727 & 0.774\\
\bottomrule
\end{tabular}}

@@

For comparison, the regression tree produces a similar prediction to the
logic regression, however, we argue that the logic regression tree shown
in Figure~\ref{fig-logic-reg} is more interpretable for our purposes
because it provides a direct representation of which combinations of
analysis checks lead to unexpected outcomes. The logic regression tree
is also directly comparable to other diagnostic techniques, which we
discuss further in Section~\ref{sec-discussion}.

@@

\phantomsection\label{cell-fig-dist-fit}
\begin{figure}[H]

@@

@@

\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}{>{\raggedright\arraybackslash}p{18em}rrrrr}
\toprule
Checks & Precision & Recall & Independence & Harmonic & Arithmetic\\
\midrule
Check 1: cor(m, PM10) < -0.03 & 0.551 & 0.894 & 1 & 0.763 & 0.815\\
Check 2: cor(m, tmp) > -0.35 & 0.281 & 0.556 & 1 & 0.472 & 0.612\\
Check 3: mortality outlier & 0.302 & 0.775 & 1 & 0.536 & 0.692\\
Check 4: PM10 outlier & 0.305 & 0.766 & 1 & 0.537 & 0.690\\
Logic regression: (check 1) AND ((check 2) OR (check 3 AND check 4)) & 0.722 & 0.876 & 1 & 0.851 & 0.866\\
\bottomrule
\end{tabular}}

@@

This generates a precision of 0.722 and a recall of 0.876 for predicting
the unexpected PM10 coefficient. As shown in
Figure~\ref{fig-linear-reg-tree}, there is no single analysis check in
the tree that predicts an unexpected outcome. rather at least three
checks in the tree must be TRUE in order for the model to predict an
unexpected outcome. Given the high independence of the checks
(Table~\ref{tbl-linear-reg}), this suggests that unexpected results are
only likely after multiple anomalies are observed in the data.

@@

@@

Systematically generating realistic simulated data is a key component of
our approach and is deserving of further consideration. In the PM10
example in Section~\ref{sec-pm10-mortality}, the inverse transform
method was used to preserve the correlation structure among mortality,
PM10, and temperature. However, the simulation process can become
complex when additional restrictions are imposed or when a greater range
of scenarios is desired. In such cases, techniques like the
acceptance-reject method or permutation may be used to generate the
data. The results in Section~\ref{sec-pm10-mortality} on the broader
NMMAPS dataset suggest that our simulation procedure may have been
inadequate in reflecting the range of possible configurations that the
data could take. In particular, our split-data approach, using New York
City to guide the simulation and then applying the analysis checks to
other cities, may not have been ideal. It may be worth exploring some
recent work in data thinning \citep{neufeld2024data} and data fission
\citep{leiner2023data} to create training datasets that are more
representative of future data.

>>>>

\begin{verbatim}
Orientation inferred to be along y-axis; override with
`position_quasirandom(orientation = 'x')`
\end{verbatim}

@@

@@

While checks may score high on predictive accuracy, they may be
correlated and less useful to provide actions to act on the analysis.
This could happen if a set of checks are all tangentially related to the
cause of the unexpected results, but none addresses the root cause. To
quantify the redundancy among the checks, we use total correlation,
\(C\), a multivariate generalization of mutual information that captures
the amount of shared information across two variables. We define the
normalized total correlation, \(C'\), as

\[C'(X_1, X_2, \cdots, X_k) = \frac{C(X_1, X_2, \cdots, X_k)}{\sum_{i= 1}^k H(X_i)} = \frac{\sum_{i= 1}^k H(X_i) - H(X_1, X_2, \cdots, X_k)}{\sum_{i= 1}^k H(X_i)}, \]
where \(H(X_i)\) is the entropy of the \(i\)-th check and
\(H(X_1, X_2, \cdots, X_k)\) is the joint entropy of checks
\(\{X_1, X_2, \cdots, X_k\}\). The amount of independent information,
\(\eta\), can be defined as:

\[\eta(X_1, X_2, \cdots, X_k) = 1 - C'(X_1, X_2, \cdots, X_k) = \frac{H(X_1, X_2, \cdots, X_k)}{\sum_{i= 1}^k H(X_i)}.\]

To scale this quantity to the range \([0,1]\) for consistency with other
metrics (i.e., precision and recall), we define the independence metric
as:

\[\text{independence} = \frac{\eta - 1/k}{1 - 1/k}= \frac{H(X_1, X_2, \cdots, X_k)/\sum_{i= 1}^K H(X_i) - 1/k}{1 - 1/k}.\]

The metric attains a value of 1 indicates that all checks are
independent: \(H(X_1, X_2, \cdots, X_k)=\sum_{i= 1}^k H(X_i)\) and
provide unique information to diagnose the unexpected outcome (trivia
case when there is only one check), and a value of 0 when the checks are
identical:
\(H(X_1, X_2, \cdots, X_k)- \sum_{i= 1}^k H(X_i) = H(X_i)/ k H(X_i) = 1/k\).

@@

@@

\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}{>{\raggedright\arraybackslash}p{19em}rrrrr}
\toprule
Checks & Precision & Recall & Independence & Harmonic & Arithmetic\\
\midrule
Check 1: q(step, 0.6) > 10000 & 0.319 & 0.575 & 1.000 & 0.511 & 0.631\\
Check 2: q(step, 0.4) < 8000 & 0.264 & 0.613 & 1.000 & 0.467 & 0.626\\
Check 3: sd(step) > 2500 & 0.153 & 0.289 & 1.000 & 0.273 & 0.481\\
Logic regression: (check 1 OR check 2) AND (not check 3) & 0.861 & 0.431 & 0.818 & 0.637 & 0.703\\
Comparison: (check 1) AND (not check 3) & 0.278 & 0.870 & 0.881 & 0.510 & 0.676\\
\addlinespace
Regression tree & 0.542 & 0.780 & 0.818 & 0.689 & 0.713\\
\bottomrule
\end{tabular}}

@@

For comparison, we include a simplified rule, (quantile(step, 0.6)
\textgreater{} 10,000) AND (NOT sd(step) \textgreater{} 2,500). Although
the checks on the 60th quantile and standard deviation are less
correlated -- reflected in a higher independence score -- this pair is
less accurate at predicting the unexpected outcome, resulting in lower
harmonic and arithmetic mean scores. The regression tree produces a
similar prediction to the logic regression, however, we argue that the
logic regression tree shown in Figure~\ref{fig-logic-reg} is more
interpretable for our purposes because it provides a direct
representation of which combinations of analysis checks lead to
unexpected outcomes. The logic regression tree is also directly
comparable to other diagnostic techniques, which we discuss further in
Section~\ref{sec-discussion}.

@@

@@

\begin{verbatim}
Warning: There were 990 warnings in `mutate()`.
The first warning was:
i In argument: `fit = list(...)`.
i In row 2.
Caused by warning in `dpois()`:
! non-integer x = 291.790133
i Run `dplyr::last_dplyr_warnings()` to see the 989 remaining warnings.
\end{verbatim}

\begin{verbatim}
Joining with `by = join_by(unexpect)`
\end{verbatim}

@@

\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}{>{\raggedright\arraybackslash}p{18em}rrrrr}
\toprule
Checks & Precision & Recall & Independence & Harmonic & Arithmetic\\
\midrule
Check 1: cor(m, PM10) < -0.03 & 0.551 & 0.894 & 1.000 & 0.763 & 0.815\\
Check 2: cor(m, tmp) > -0.35 & 0.281 & 0.556 & 1.000 & 0.472 & 0.612\\
Check 3: mortality outlier & 0.302 & 0.775 & 1.000 & 0.536 & 0.692\\
Check 4: PM10 outlier & 0.305 & 0.766 & 1.000 & 0.537 & 0.690\\
Logic regression: (check 1) AND ((check 2) OR (check 3 AND check 4)) & 0.876 & 0.722 & 0.812 & 0.798 & 0.803\\
\bottomrule
\end{tabular}}

@@

This generates a precision of 0.876 and a recall of 0.722 for predicting
the unexpected PM10 coefficient. As shown in
Figure~\ref{fig-linear-reg-tree}, there is no single analysis check in
the tree that predicts an unexpected outcome. rather at least three
checks in the tree must be TRUE in order for the model to predict an
unexpected outcome. Given the high independence of the checks
(Table~\ref{tbl-linear-reg}), this suggests that unexpected results are
only likely after multiple anomalies are observed in the data.

@@

\begin{verbatim}
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 3288 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr  (1): season
dbl  (4): mortality, pm10, temp, dewpt
date (1): date

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Rows: 40 Columns: 2
-- Column specification --------------------------------------------------------
Delimiter: ","
chr (2): city, cityname

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
Joining with `by = join_by(city)`
\end{verbatim}

@@

Systematically generating realistic simulated data is a key component of
our approach and is deserving of further consideration. In the PM10
example in Section~\ref{sec-pm10-mortality}, the inverse transform
method was used to preserve the correlation structure among mortality,
PM10, and temperature. However, the simulation process can become
complex when additional restrictions are imposed or when a greater range
of scenarios is desired. In such cases, techniques like the
acceptance-reject method or permutation may be used to generate the
data. The results in Section~\ref{sec-pm10-mortality} on the broader
NMMAPS dataset suggest that our simulation procedure may have been
inadequate in reflecting the range of possible configurations that the
data could take. In particular, our split-data approach, using New York
City to guide the simulation and then applying the analysis checks to
other cities, may not have been ideal. It may be worth exploring some
recent work in data thinning \citep{neufeld2024data}, data fission
\citep{leiner2023data}, or differential privacy methods
\citep{dong_gaussian_2022} to create training datasets that are more
representative of future data. The Gaussian copula used in the example
doesn't account for the tail dependence between the variables and other
copula, e.g.~t-copula or extreme value distribution copula, could also
be explored to better capture the tail dependence.
